Demystifying the Combination of Dynamic Slicing and \\ Spectrum-based Fault Localization

### 2019-02-24 paper/ijcai/paper.tex
cleanup
more fixes

As to quantify complexity of fault localization on this dataset, we
additionally counted the number of cases of ``fault of omission'', a
popular metric in fault localization research for that
purpose~\cite{Pearson:2017:EIF:3097368.3097441,just-defects4j-issta2014}. 
A fault of omission occurs when
the program is buggy because some code is missing, i.e., a bug that is
corrected by adding code. Intuitivelly, the ability of techniques to
produce accurate diagnostic reports in those cases is poor since fault of omission
are non-executable statements which means that the execution information
cannot help locate the fault directly. A total of $59.7\%$ of the cases
in this dataset manifest fault of omission, which is a good indication of
difficulty\Sof{icse17 check}.

%in producing
%high-qualitiy diagnosis.

@@

\subsection{Results}

This section answers the research questions.

@@

@@

A fundamental issue of purely dynamic slicing techniques is the risk
of discarding faulty statements (see Section~\ref{sec:slicing}). This
research question evaluates the practical impact of this conceptual
issue. Table~\ref{tab:ds-cases-captured} shows the number of cases
dynamic slicing captures at least one of the faulty statements among
the sliced files for different values of $k$. These numbers indicate
that in general \ds{} finds at least one faulty statement in the top
$5$ of the highest-ranked faulty classes ($87.3\%$ on average) which
is slightly improved for $91.2\%$ if we consider
the top $10$ of the highest-ranked faulty classes.

@@

		\bottomrule
	\end{tabular}
	\caption {\ds{} performance on capturing faulty statements, as
     percentages. Higher is better.}
	\label{tab:red}
   \label{tab:ds-cases-captured}
\end{table}
\normalsize

@@

To analyze the impact of dynamic slicing in context, we also compared
\combpar{k} and \sfl{}$^{k}$ considering the number of cases where the
fault is captured by the technique. Note that the reason \sfl{}$^{k}$
misses the fault is different compared to \combpar{k}. \sfl{}$^{k}$
captures the fault if the fault is included within the top $k$ files;
it misses the fault otherwise. Table~\ref{table:fsws} shows results,
indicating that \combpar{k}, on average, captures more faults compared
to \sfl{}$^{k}$. In only two cases \sfl{}$^{k}$ outperformed
\combpar{k}. These results show that the majority of faulty statements
are found on the $5$ highest ranked faulty classes and that
\combpar{10} is the variant of our approach that performs better
missing only 23 out of the \numFaults\ ($\sim$8.8\%) faults.

@@

\begin{table}[h!]
	\small
	\setlength{\tabcolsep}{3pt}
	\centering
\begin{tabular}{l|rr|rr}
		 \toprule
       & \multicolumn{2}{c|}{$k=5$} & \multicolumn{2}{c}{$k=10$} \\
		 Project        &   \sfl{}   & \comb &   \sfl{}   & \comb\\
		\midrule
		 \lang{}         &   84.6   & 100.0 & 84.6 & 96.9  \\
		\cmath{}           & 81.7 & 89.4 & 85.6 & 95.2 \\
		\chart{}           & {\cellcolor{Gray} 84.6} & {\cellcolor{Gray}76.9} & {\cellcolor{Gray}92.3} & {\cellcolor{Gray}84.6}\\
		\jtime{}           & 77.8 & 81.5 & 81.5 & 85.2\\
		\mockito{}           & 63.2 & 71.1 & 71.1 & 78.9\\   \midrule
		Total         & 79.6 & 87.3 & 83.5 & 91.2\\
		\bottomrule
	\end{tabular}
  \caption {Number of faults where at least one of the faulty-statements is in the \sfl{}$^{k}$ and \combpar{k} report}
  \label{table:fsws}
\end{table}

@@

In \textbf{RQ3}, we evaluate the effective impact of combining \ds{}
with \sfl{} to improve fault localization.

The first experiment we ran to assess impact consists of measuring the
\emph{difference of diagnosis cost}. This metric has been previously
used in other studies~\Fix{cite cite}. More specifically, we computed
$\Delta{}C = C(\textrm{SFL$^{k}$}) - C(\textrm{\combpar{k}})$, where
$C$ denotes the diagnosis cost and is obtained by computing the mean
position in the ranking of all buggy statements. $\Delta C >0$ means
that the position of the faulty statement has decreased with
\combpar{k}, \ie{}, the combination had a positive impact. $\Delta C
<0$ means that \combpar{k} performs worse compared to its
baseline. That could happen if dynamic slicing misses the faulty
statement---the statements not included in the slice will have their
suspiciousness scores reduced and ranks potentially reduced. $\Delta
C=0$ means that the faulty statement remained in the same position.
It is worth noting that we assume a perfect bug understanding, i.e.,
if the developer sees the bug in the ranking, he is able to precisely
determine if it is the faulty one.

Figure \ref{fig:diagnosis} summarizes performance of \combpar{k}
relative to \sfl{}$^{k}$ using histograms over $\Delta{}C$ for
different values of $k$. Note that (1) the number of cases where
\sfl{}$^{k}$ outperforms \combpar{k} is very small compared and (2)
the number of cases \combpar{k} outperforms \sfl{}$^{k}$ is
significant and, as we increase the value of $k$, it increases (from
$97$ to $117$ faults).

@@

\begin{figure}[h]
		\centering
		\includegraphics[width=0.42\textwidth]{figures/performance.pdf}
		\caption{Histograms of $\Delta{}C$.}
		\label{fig:diagnosis}
\end{figure}

\begin{figure}[h]
		\centering
		\includegraphics[width=0.49\textwidth]{figures/boxplot.pdf}
		\caption{Boxplot of $\Delta{}C$.}
		\label{fig:boxplot}
\end{figure}
%
% \begin{figure}[h]
% 		\centering
% 		\includegraphics[width=0.42\textwidth]{figures/boxplot_all.pdf}
% 		\caption{Boxplot of $\Delta{}C$. \Sof{Considering the best cases and the \textbf{top-misses}}}
% 		\label{fig:boxplot}
% \end{figure}

If we focus at the $116$\Mar{117?} faults where there was improvement
in at least one of the variants, the mean improvement in the ranking
was 41.1 statements for $k=5$ and 18.7 statements for $k=10$. We
observed an impressive maximum ranking improvement of $3170.5$\Mar{why
  the .5 if this is the maximum case?} statements for $k=5$ and $209$
statements for $k=10$. Such difference is justified by the fact that
entire files are discarded in \combpar{5}, whose statements are
included in the ranking of \combpar{10}. Consequently, the
improvements obtained by the technique will not be as high if slicing
on these additional files is unable to aggressively and safely discard
statements. This also shows that slicing files, in addition to slicing
statements from those files, is very beneficial for fault localization
performance.

\Mar{show distribution...}

In the following, we present statistical analysis to demonstrate that
the distributions of $C$ for each technique are different.
Table~\ref{table:st}, reports the average rank of all faulty
statements of the \dfj{} faults considered to the study after applying both
techniques, \sfl{}$^{k}$ and \combpar{k} for $k=5$ and $k=10$. $33$ of the $260$
faulty versions are cases whose faulty statements decreased in the ranking list due to being a fault of omission case or due to limitations regarding the test design which \ds{} could not handle. Thus, we report both values in Table~\ref{table:st}. In bold, we present the
results of the $227$ faults that do not contain the cases that are result of the limitations mentioned before. Inside parentheses, we present the
rank results for all faults. The \combpar{k} combination yields an improvement
of $17.3$ on the rank mean of the dataset faulty statements when using $k=5$.
Instead, if $k=10$ is considered the proposed approach returns an improvement of
$6.1$ on the rank mean of all faulty statements.
These $227$ faults contain cases where the faulty class was not in the choosen set of classes $k$ and a high amount of cases where the improvement was none.

\begin{table}[h!]
	\tiny
	\centering
	\setlength{\tabcolsep}{3pt}
	\begin{tabular}{@{}c|c|c|c|c@{}}
\toprule
  & SFL$^{5}$      & SFL$^{10}$     & \combpar{5}              & \combpar{10}              \\ \midrule
Mean & \textbf{590.7} (584.3)   &  \textbf{512.7} (516.4)  & \textbf{573.4} (679.7)  & \textbf{506.6} (614.0)   \\ \midrule
Median & \textbf{46.5} (49.4)    & \textbf{54.5} (53.0) & \textbf{42.7} (69.3)  & \textbf{40.0} (56.9)\\ \midrule
Variance & \textbf{1257.0} (1223.1)  &  \textbf{1171.5} (1148.1) & \textbf{1239.1} (1352.4) &  \textbf{1167.0} (1294.9) \\ \midrule
Shapiro-Wilk &\makecell{$W$ = 0.52 \\ $p$ = 1.75 x $10^{-24}$} & \makecell{$W$ = 0.48 \\ $p$ = 2.50 x $10^{-25}$} & \makecell{$W$ = 0.51 \\ $p$ = 1.10 x $10^{-24}$} & \makecell{$W$ = 0.48 \\ $p$ = 2.26 x $10^{-25}$}  \\ \midrule
Friedman & \multicolumn{4}{c}{\makecell{$\chi^{2}$ = 162.61 \\ p-value = 5.00 x $10^{-35}$}} \\
\bottomrule
\end{tabular}
  \caption {Statistical tests}
  \label{table:st}
\end{table}

@@

To understand which techniques perform differently, \ie{} to answer
the question \textit{Does \combpar{k} perform differently than
  \sfl{}$^{k}$}, we performed a Siegel post-hoc analysis. Figure
\ref{fig:performance} reports the results of Siegel analysis for all
pairs of techniques. Each square\Mar{gray intensity?} represents the
significance\Mar{of what?} between the corresponding techniques, i.e.,
if it is significant it means that the techniques perform
differently. We can determine with $95\%$ confidence that the
performance of each pair of techniques is different.

>>>>

As to quantify complexity of fault localization on this dataset, we
additionally counted the number of cases of ``fault of omission'', a
popular metric in fault localization research for that
purpose~\cite{Pearson:2017:EIF:3097368.3097441}. 
A fault of omission (\foo) occurs when
the program is buggy because some code is missing, i.e., a bug that is
corrected by adding code. Intuitively, the ability of techniques to
produce accurate diagnostic reports in those cases might be poor since 
fault of omission bugs are very difficult to locate. This dataset has a
total of $59.7\%$ of \foo s which may contribute negatively to our
results since \sfl{} strives to rank these cases.
%in producing
%high-qualitiy diagnosis.

@@

\subsection{Results and Discussion}

This section discusses and answers the research questions.

@@

\begin{table}[H]
  \small
	\centering
	\setlength{\tabcolsep}{4pt}
	\begin{tabular}{lrr}
		\toprule
		Project             &  \multicolumn{1}{c}{$k=5$} & \multicolumn{1}{c}{$k=10$} \\ %\combpar{5}  & \combpar{10}
		\midrule

@@

A fundamental issue of purely dynamic slicing techniques is the risk
of discarding faulty statements (see Section~\ref{sec:slicing}). This
research question evaluates the practical impact of this conceptual
issue. Table~\ref{tab:ds-cases-captured} shows the number of cases
dynamic slicing captures at least one of the faulty statements among
the sliced files for different values of $k$. These numbers indicate
that in general \ds{} finds at least one faulty statement in the top
$5$ of the highest-ranked faulty classes ($87.3\%$ on average) which
is slightly improved to $91.2\%$ if we consider
the top $10$ of the highest-ranked faulty classes.

@@

		\bottomrule
	\end{tabular}
	\caption {\ds{} performance on capturing faulty statements, as
     percentages. Higher is better.}
   \label{tab:ds-cases-captured}
\vspace{-0.2cm}

\end{table}
%\normalsize

@@

To analyze the impact of dynamic slicing in context, we also compared
\combpar{k} and \sfl{}$^{k}$ considering the number of cases where the
fault is captured by the technique. Note that the reason \sfl{}$^{k}$
misses the fault is different compared to \combpar{k}. \sfl{}$^{k}$
captures the fault if the fault is included within the top $k$ files;
it misses the fault otherwise. Table~\ref{table:fsws} shows results,
indicating that \combpar{k}, on average, captures more faults compared
to \sfl{}$^{k}$. In only two cases \sfl{}$^{k}$ outperformed
\combpar{k}. These results show that the majority of faulty statements
are found on the $5$ highest ranked faulty classes and that
\combpar{10} is the variant of our approach that performs better
missing only $23$ out of the \numFaults\ ($\sim$8.8\%) faults.

@@

\begin{table}[h!]
	\vspace{-0.2cm}

	\small
	\setlength{\tabcolsep}{3pt}
	\centering
\begin{tabular}{l|rr|rr}
		 \toprule
       & \multicolumn{2}{c|}{$k=5$} & \multicolumn{2}{c}{$k=10$} \\
		 Project        &   \sfl{}   & \comb &   \sfl{}   & \comb\\
		\midrule
		 \lang{}         &   84.6   & 100.0 & 84.6 & 96.9  \\
		\cmath{}           & 81.7 & 89.4 & 85.6 & 95.2 \\
		\chart{}           & {\cellcolor{Gray} 84.6} & {\cellcolor{Gray}76.9} & {\cellcolor{Gray}92.3} & {\cellcolor{Gray}84.6}\\
		\jtime{}           & 77.8 & 81.5 & 81.5 & 85.2\\
		\mockito{}           & 63.2 & 71.1 & 71.1 & 78.9\\   \midrule
		Total         & 79.6 & 87.3 & 83.5 & 91.2\\
		\bottomrule
	\end{tabular}
  \caption {Number of faults where at least one of the faulty-statements is in the \sfl{}$^{k}$ and \combpar{k} report}
  \label{table:fsws}
\vspace{-0.4cm}
\end{table}

@@

In \textbf{RQ3}, we evaluate the effective impact of combining \ds{}
with \sfl{} to improve fault localization. The first experiment we ran to assess impact consists of measuring the
\emph{difference of diagnosis cost}. This metric has been previously
used in other studies~\cite{7390282,ang-perez-van-deursen-rui-2017,Pearson:2017:EIF:3097368.3097441,Perez:2018:LQR:3304889.3304927}. More specifically, we computed
$\Delta{}C = C(\textrm{SFL$^{k}$}) - C(\textrm{\combpar{k}})$, where
$C$ denotes the diagnosis cost and is obtained by computing the mean
position in the ranking of all buggy statements. $\Delta C >0$ means
that the position of the faulty statement has decreased with
\combpar{k}, \ie{}, the combination had a positive impact. $\Delta C
<0$ means that \combpar{k} performs worse compared to its
baseline. That could happen if dynamic slicing misses the faulty
statement---the statements not included in the slice will have their
suspiciousness scores reduced and ranks potentially reduced. $\Delta
C=0$ means that the faulty statement remained in the same position.
It is worth noting that we assume a perfect bug understanding, i.e.,
if the developer sees the bug in the ranking, he is able to precisely
determine if it is the faulty one.

Figure \ref{fig:diagnosis} summarizes performance of \combpar{k}
relative to \sfl{}$^{k}$ using histograms over $\Delta{}C$ for
different values of $k$. Note that (1) the number of cases where
\sfl{}$^{k}$ outperforms \combpar{k} is very small compared and (2)
the number of cases \combpar{k} outperforms \sfl{}$^{k}$ is
significant and, as we increase the value of $k$, it increases (from
$97$ to $117$ faults). The faults where the cost of diagnosis is
negative are cases that (1) the faulty statements decreased in the ranking 
list due to being a fault of omission, (2) the test has design limitations
which \ds~is not capable of handling, or (3) the buggy-class in not in
the top $k$ of the highest-ranked faulty classes under analysis.

@@

\begin{figure}[h]
	\vspace{-0.3cm}
		\centering
		\includegraphics[width=0.415\textwidth]{figures/performance.pdf}
		\vspace{-0.4cm}
		\caption{Cost of Diagnosis ($\Delta{}C$) per variant}
		\label{fig:diagnosis}
\end{figure}

Figure~\ref{fig:boxplot} shows the $\Delta{}C$ distribution for each
variant in two different cases. On the left, we do not consider the
test engineering problems and faults of omission ($227$ faults). The two negative 
outliers present the cases where the buggy-class was not in the top
$k$. On the right, we only consider the cases where at least one 
of the variants resulted in an increase of $\Delta{}C$ ($116$ faults), i.e., the 
cases where \ds{} improves \sfl. Although, we observe
a considerable amount of cases where $\Delta{}C$ is negative or zero in Figure~\ref{fig:diagnosis}, we also see that there is a considerable amount of faults which could benefit a lot with \combpar{k}, in Figure~\ref{fig:boxplot}.
\begin{figure}[h]
		\vspace{-0.2cm}
		\centering
		\includegraphics[width=0.49\textwidth]{figures/boxplot.pdf}
		\vspace{-0.6cm}

		\caption{Distribution of $\Delta{}C$ per variant when not considering 
		test engineering problems and fault of omissions (left) and only considering
		the cases where \combpar{k} achieved improvement under one of the variants}
		\label{fig:boxplot}
		\vspace{-0.2cm}

\end{figure}

If we consider only the faults which can benefit from our approach, we can observe that there is a considerable
amount of cases where the approach is actually successful. For
$k=5$, our approach achieves a mean improvement of $41.1$ ($9.57\%$ on average) statements
on the ranking and an impressive maximum ranking improvement
of $3171$ ($96.7\%$ on average) statements. Whereas for $k=10$, \combpar{k}
achieves a mean improvement of $19$ ($13.35\%$ on average) statements
on the ranking and a maximum ranking improvement
of $209$ ($73.7\%$ on average) statements. Such difference is justified by the fact that
entire files are discarded in \combpar{5}, whose statements are
included in the ranking of \combpar{10}. Consequently, the
improvements obtained by the technique will not be as high if slicing
on these additional files is unable to aggressively and safely discard
statements. This also shows that slicing files, in addition to slicing
statements from those files, is very beneficial for fault localization
performance.
In the following, we present statistical analysis to demonstrate that
the distributions of $C$ for each technique are different.
Table~\ref{table:st}, reports the average rank of all faulty
statements of the \dfj{} faults considered to the study after applying both
techniques, \sfl{}$^{k}$ and \combpar{k} for $k=5$ and $k=10$. 
In bold, we present the results of the $227$ faults that do not contain the 
cases that are result of the limitations mentioned before (test design and fault of omission). Inside parentheses, we present the
rank results for all faults. The \combpar{k} combination yields an improvement
of $17.3$ on the rank mean of the dataset faulty statements when using $k=5$.
Instead, if $k=10$ is considered, the proposed approach returns an improvement of
$6.1$ on the rank mean of all faulty statements.
These $227$ faults contain cases where the faulty class was not in the choosen 
set of classes $k$ and a high amount of cases where the improvement was none.

\begin{table}[h!]
	\tiny
	\centering
	\setlength{\tabcolsep}{3pt}
	\begin{tabular}{@{}c|c|c|c|c@{}}
\toprule
  & SFL$^{5}$      & SFL$^{10}$     & \combpar{5}              & \combpar{10}              \\ \midrule
Mean  & \textbf{590.7} (584.3)   &  \textbf{512.7} (516.4)  & \textbf{573.4} (679.7)  & \textbf{506.6} (614.0)   \\ \midrule
Median & \textbf{46.5} (49.4)    & \textbf{54.5} (53.0) & \textbf{42.7} (69.3)  & \textbf{40.0} (56.9)\\ \midrule
Variance & \textbf{1257.0} (1223.1)  &  \textbf{1171.5} (1148.1) & \textbf{1239.1} (1352.4) &  \textbf{1167.0} (1294.9) \\ \midrule
Shapiro-Wilk &\makecell{$W$ = 0.52 \\ $p$ = 1.75 x $10^{-24}$} & \makecell{$W$ = 0.48 \\ $p$ = 2.50 x $10^{-25}$} & \makecell{$W$ = 0.51 \\ $p$ = 1.10 x $10^{-24}$} & \makecell{$W$ = 0.48 \\ $p$ = 2.26 x $10^{-25}$}  \\ \midrule
Friedman & \multicolumn{4}{c}{\makecell{$\chi^{2}$ = 162.61 \\ p-value = 5.00 x $10^{-35}$}} \\
\bottomrule
\end{tabular}
  \caption {Statistical tests for $C$}
  \label{table:st}
\end{table}

@@

To understand which techniques perform differently, \ie{} to answer
the question \textit{Does \combpar{k} perform differently than
  \sfl{}$^{k}$}, we performed a Siegel post-hoc analysis. Figure
\ref{fig:performance} reports the results of Siegel analysis for all
pairs of techniques. Each square represents the
statistical significance of the diagnostic accuracy amongst the different techniques, i.e.,
if it is significant it means that the techniques perform
differently. We can determine with $95\%$ confidence that the
performance of each pair of techniques is different.
